---
title: "Reproducible Research: Peer Assignment 1"
output: 
  html_document: 
    keep_md: yes
---


## Loading and preprocessing the data 

1. Load the data

```{r Assignment1-1, echo=TRUE, fig.path="figure"}

  library(knitr)
  opts_knit$set(root.dir=normalizePath('./'))
  opts_chunk$set(fig.path = "./figures/")

  ## language setting for different native language.
  Sys.setlocale(category = "LC_ALL", locale = "english")

  # set working dir
  ## setwd(paste(getwd(),"/PA1",sep=""))

  ## read CSV file
  data<-read.csv("./data/activity.csv")
  

  # check the variables
  str(data)

  # check the dataset size(dimension)
  dim(data)
  
  
```

2. Process/transform the data (if necessary) into a format suitable for your analysis.

```{r Assignment1-2,echo=TRUE}


  # remove NAs
  data.noNA <- data.no.miss.by.date<-data[complete.cases(data),]
  
  # generate data for histogram used to answering first question.
  library(dplyr)
  library(tidyr)
  
  data.no.miss.by.date<-data.noNA %>% group_by(date) %>% summarize(step.sum=sum(steps))
  
  ## check the data.
  head(data.no.miss.by.date)

```


## What is mean total number of steps taken per day.


1. Make a histogram of the total number of steps taken each day

```{r Assignment2-1,echo=TRUE}
  require(ggplot2)
  p<-ggplot(data.no.miss.by.date,aes(date,step.sum)) + geom_bar(stat="identity",width=1) + theme_bw(base_size=10)
  p<-p+ theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5)) + ylab("total steps per a day")
  p 
  
```

2. What  is mean total number of steps taken per day

```{r Assignment2-2,echo=TRUE}
  ## calculate the mean and median.
  data.mean<-mean(data.no.miss.by.date$step.sum)
  data.median<-median(data.no.miss.by.date$step.sum)
```

  - the mean of steps taken per a day is `r data.mean`.
  - the median of steps taken per a day is `r data.median`.

## What is the average daily activity pattern ?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r Assignment3-1, echo=TRUE}
   data.no.miss.by.interval<-data[complete.cases(data),] %>% group_by(interval) %>% summarize(mean=mean(steps))
  
   p<-ggplot(data.no.miss.by.interval,aes(interval,mean)) + geom_line() + theme_bw() + xlab("5-minute interval") + ylab("Average number of steps taken") + geom_vline(xintercept=835, linetype=2,color="red")
   p
  
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r Assignment3-2,echo=TRUE}
  max.step.interval<-head(data.no.miss.by.interval[order(-data.no.miss.by.interval$mean),]$interval,1)
  
```

  - the time interval of `r max.step.interval` is the time interval containing maximum number of steps.

## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r Assignment4-1,echo=TRUE}
  nacols<-is.na(data$steps)
  sum(nacols)
  
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r Assignment4-2,echo=TRUE}
  # get the number of intervals
  numInterval<-dim(data.no.miss.by.interval)[1]
  
  new.impute.mean <- data
  # get rownames(index) of the NA values.
  
  na.index<-as.numeric(rownames(new.impute.mean[nacols,]))%%numInterval
  na.index<-ifelse(na.index==0,numInterval,na.index)
  
  # compare the length of two vector
  
  length(na.index)==sum(nacols)
  
  # now we can impute missing values using na.index. 
  
 
   
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r Assignment4-3,echo=TRUE}
  new.impute.mean[nacols,]$steps <- data.no.miss.by.interval[na.index,]$mean
  
  # check missing
  sum(is.na(new.impute.mean))
  
```
  
  - new data has 'r sum(is.na(new.impute.mean))' NAs

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r Assignment4-4,echo=TRUE}
  
  data.no.miss.by.date.impute<-new.impute.mean %>% group_by(date) %>% summarize(step.sum=sum(steps))

  require(ggplot2)
  p<-ggplot(data.no.miss.by.date.impute,aes(date,step.sum)) + geom_bar(stat="identity",width=1) + theme_bw(base_size=10)
  p<-p+ theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=.5)) + xlab("total steps taken")
  p
  
  ## calculate the mean and median.
  data.mean<-mean(data.no.miss.by.date.impute$step.sum)
  data.median<-median(data.no.miss.by.date.impute$step.sum)
```

  - the mean of steps taken per a day is `r data.mean`.
  - the median of steps taken per a day is `r data.median`.
  


## Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

  1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r Assignment5-1,echo=TRUE}
  new.impute.mean$WKDAY<-weekdays(as.Date(new.impute.mean$date))
  SatSun<-c("Saturday","Sunday")
  
  
  imputed.data <- new.impute.mean %>% mutate(WKDAY=as.factor(ifelse(WKDAY %in% SatSun,"Weekend","Weekday")))
  
  head(imputed.data)
  
  
  
      
```

  2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r Assignment5-2,echo=TRUE}
  imputed.data.summarized<- imputed.data %>% group_by(interval,WKDAY) %>% summarize(mean=mean(steps))
  
  require(ggplot2)
  p<-ggplot(imputed.data.summarized,aes(interval,mean)) + geom_line() + theme_bw() + xlab("5-minute interval") + ylab("Average number of steps taken") + facet_wrap(~ WKDAY,ncol=1,nrow=2)
   p
  
```
